---
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
params:
  origin_login_file: Malawi
  cache: TRUE
  echo: FALSE
---

<!-- NB:  when changing to country for review, MUST clear cache.  If not, markdown throws error: 'Error: path for html_dependency not found:' -->

# **Desktop Review: `r toupper( params$origin_login_file )`**


```{r packages, echo = TRUE, cache=FALSE, warning=FALSE, include = FALSE }


# list of required packages
    package.list = c("rlist", "listviewer", "knitr", "kableExtra", "xts", "leaflet", "RColorBrewer", "DT", "dygraphs", "httr", "jsonlite", "XML", "assertthat", "lubridate", "scales", "RcppRoll", "zoo", "gridExtra", "futile.logger", "tidyverse" )

# Function to test if package is installed 
    pkgTest <- function( package.list = package.list ){
        
        missing.packages = setdiff( package.list , rownames(installed.packages())) 
        if ( length( missing.packages ) > 0 ) install.packages( missing.packages ) 
    }


# Test if packages loaded
    pkgTest( package.list )

# load the packages
    lapply( package.list, suppressMessages( require ) , character.only = TRUE)


knitr::opts_chunk$set( echo = params$echo ,
                       fig.width =  8, 
                       # knitr.table.format = "html" ,
                       cache = params$cache # TRUE for testing; FALSE for production
                       )

```


```{r sources, cache = FALSE}

  # functions for fetching metadata

  source( 'dhis2_functions.R'  )
  
  # folder for storing review data

  dhis2 = "../HMIS/DHIS2/"  # Location of base files
  # dhis2 = ""  # Location of base files
 
  folder = paste0( dhis2 ,  params$origin_login_file  , "/" )
    
  if ( !dir.exists( folder ) ){
    dir.create( folder ) 
  } 
 
  
  # login shortcut
   login = function(){
       source( paste0( folder, 
                       tolower( params$origin_login_file ) , 
                       "_login")  )
       loginDHIS2( baseurl, username, password)
   }
 

```

<!-- Define the year-month time periods for the review -->

<!-- NB:  What about IDSR with weeks?  -->

```{r date_strings }

 # Function to create string of dates
  date_code = function( years = 2013:2017 , months = 1:12 ){
    period = character()
    
    for (year in years ){
      for (month in seq_along(months)){
        this_period = paste0( ";", 
                         year , 
                         ifelse( month < 10 , paste0("0", month) , month )
        )
        period =  c(period, this_period )
      }
    }
    
    # remove first ;
    period =  paste( period, collapse = "")
    period = substring( period, 2, nchar(period))
    
    return( period )
  }
 
  periods = date_code() # default: 2013-2017
  # periods = date_code( 2017, 1 ) # January 2017 only
  
  # as character vector
  periods.vector = strsplit( periods, ";" , fixed = TRUE )[[1]]

  
```



# Metadata 


```{r all_meta }

  meta_data_file = paste0( folder , params$origin_login_file, 
                           "_metadata.rds" ) 

  if ( !file.exists( meta_data_file) ){
   
    # login in to server
    login()
  
    md = metadataDHIS2( baseurl, 'all' )
    
    # glimpse( md )
    
    write_rds( md, meta_data_file )
    
  } else {
      
       md = read_rds( meta_data_file )
  }

  # re-order alaphabetically to be easier to browse in View
  md = md[order(names(md))]
  
  # Retrieve access data 
  date_metadata = file.info( meta_data_file )$ctime 
  
```

Results shown are for data accessed on **`r format(date_metadata, "%d %B %Y") `**.    

`r params$origin_login_file` is running DHSI2, version **`r md$system$version`** (rev. `r md$system$rev`), last updated `r md$system$date`.

```{r list_viewer }
 jsonedit( md )
```

## Numbers of each attribute

```{r attribute_numbers }

 metadata_attributes = map_df( md[names(md)], ~length(.x$id) ) %>% 
    gather( attribute, value ) %>% # columns to rows
    mutate( value = comma( value ) )   # format numbers
    

  metadata_attributes %>%
      kable( align = c( 'l', 'r' ), 'html' ) %>% 
      kable_styling( bootstrap_options = c("striped", "hover") ) %>%
        column_spec(1, bold = T) %>%
      scroll_box(height = "500px")

```

## Organizational Units and levels

```{r ou_and_levels }

feature_type = function( coordinate ){
    n = length( gregexpr( '[' , coordinate , fixed = TRUE)[[1]] )
    
    if ( is.na( coordinate ) ) return(NA)
    if (n==1) return('Point') 
    if (n>1) return('Polygon') 
    

}

  # levels
  levels = character()
      for ( i in seq_along( md$organisationUnitLevels$id) ){
    
        levels =  c(levels, paste0("LEVEL-", i) )
      }
      
      levels = paste( levels, collapse = ";")
  
  levels.vector = strsplit( levels, ";" , fixed = TRUE )[[1]]


   ous =  md$organisationUnits %>% 
        select( id, path, name, shortName, coordinates , 
                created, lastUpdated, openingDate, closedDate 
                ) %>% as.tibble %>%
        mutate( 
          level = map_int( path, 
                               ~length( gregexpr("/", .x, perl = TRUE)[[1]]) 
                               ) ,
          feature = map_chr( coordinates, ~feature_type(.x) )
        )
    

  table_ou_levels =   count( ous, level ) %>%
      # left_join( select( md$organisationUnitLevel, offlineLevels, name) ,
      left_join( select( md$organisationUnitLevel, name, level ) ,
                  # by = c('level'='offlineLevels') ) %>% 
                  by = c('level'='level') ) %>% 
            filter( !grepl( 'Level', name )  ) %>% # removes generic levels ('Level-8')
      select( level, name, n ) 
  
  
  # Add feature types
  ousFeatures = count( ous, level, feature ) %>% 
      spread(  feature, n )
  
  feature_table = left_join( table_ou_levels, 
                            ousFeatures, by = 'level' ) %>%
      mutate( n = comma(n) ) 
    
  kable( feature_table, align = c( 'l', 'l', 'r' ) ) 

```


There, `r sum( map_lgl( md$organisationUnits$attributeValues , ~length(.x) > 0 ) )` have an attribute values assigned to OU.  


## OU Levels

```{r ou_levels}


  # levels
      levels = character()
      for ( i in seq_along( md$organisationUnitLevels$id) ){
    
        levels =  c(levels, paste0("LEVEL-", i) )
      }
      
      levels = paste( levels, collapse = ";")
  
  # import month by month
  levels.vector = strsplit( levels, ";" , fixed = TRUE )[[1]]
  
```

summarise enumerated features by level

NB : stratify by attributes such as private, public--if available

```{r ous_by_level}


 count(ous, nchar( openingDate)>0 ) # all have

 table( year( ymd_hms( ous$openingDate )) ) # but some are recent

  
  
  n_distinct( ous$name )

  p = strsplit( periods , ";" , fixed = TRUE )[[1]]
  p = paste0( p , "01")
  p = ymd( p )
  
  # number of units per level 
   n.level = list()
   for ( i in seq_along( levels.vector ) ){
      
      n = map_int( p, ~ous %>%
                      filter( level %in% i  ) %>%
                      summarise(
                        n = sum( openingDate < .x & 
                                   ifelse( !is.na(closedDate), 
                                           closedDate > .x, TRUE )
                                 ) ) %>% .$n
      )
      n.level[[i]] = n
   }
   
  # combine into df
  ous.n = data_frame(
    level = rep( seq_along( levels.vector ) , each = length(p) ) ,
    period =  rep( periods.vector, length( levels.vector ) ) ,
    n = unlist( n.level )
  ) %>%
    filter( n > 1 ) 
  
  # convert month-Yr to date
  d = ous.n %>%
      # level names
      left_join( 
          select( md$organisationUnitLevels, level, name  ) , 
          by = 'level' ) %>%
    mutate( 
        name = fct_reorder( name, level )
        )

  d$date = fast_strptime( as.character(  d$period ) , "%Y%m") 
  
  ggplot( d,  aes(x = date, y = n, group = level )) + 
      geom_line() +
      facet_wrap( ~name , scales = 'free' )
  
  
  ###NB Join wiht ou levels for labels
  
```

NB: not all ou will be assigned to a dataset, and some may be duplicates...
## OU unit groups and group sets ?   

  - private / public
  - elimination / control 
  
NB  confirm whether all units assigned to group

## Categories linked with orgUnits
  
NB  confirm whether all units assigned to category


## Data elements and categories

NB: Focus on data elements for malaria, patients, ...


```{r}
    de = md$dataElements
    
    # Convert category combo from df to list (ends up with id as character)
    de[, 'categoryCombo' ] = as.list( de[, 'categoryCombo' ] )
    
    # dataElement vars
    # names( de )
    display.vars = c('name', 'zeroIsSignificant', 'id', 'lastUpdated',
                                 'categoryCombo' )
    
    mal = grepl( 'malaria' , de$name , ignore.case = TRUE )
    conf = grepl( 'conf' , de$name , ignore.case = TRUE )
    susp = grepl( 'susp' , de$name , ignore.case = TRUE )
    patients = grepl( 'patient' , de$name , ignore.case = TRUE )

# any malaria element
    de.mal = de[ mal, display.vars ] %>%
        rename( categoryCombo.id = categoryCombo ) %>%
        mutate( lastUpdated = format( ymd_hms( lastUpdated ), '%b %Y' ) ) %>%
        
        # categoryCombos
        left_join(  
            select( md$categoryCombos, id, name, categories ) %>%
                        rename(categoryCombo.id = id,
                               categoryCombo = name ,
                               category.id = categories
                        ) , 
                    by = 'categoryCombo.id'
                    ) %>%
    
        # dataElementGroups
        left_join( 
            md$dataElementGroups %>% 
                select( name, dataElements) %>%
                rename( Group = name ) %>% 
                unnest ,
            by = 'id'
            )    
    
    kable(
        select( de.mal , 
                name, Group, zeroIsSignificant, lastUpdated, categoryCombo
                ) %>%
            rename( dataElement = name ) %>%
            arrange( Group , dataElement), 
           'html' ,
        caption = "Confirmed malaria data elements" )  %>%
        kable_styling(bootstrap_options = c("striped", "hover")) %>%
        column_spec(1, bold = T)

    
# # confirmed malaria elements ####   
#     mal.conf = de[ mal & conf, display.vars ] %>%
#         rename( categoryCombo.id = categoryCombo ) %>%
#         mutate( lastUpdated = format( ymd_hms( lastUpdated ), '%b %Y' ) ) %>%
#         
#         # categoryCombos
#         left_join(  
#             select( md$categoryCombos, id, name, categories ) %>%
#                         rename(categoryCombo.id = id,
#                                categoryCombo = name ,
#                                category.id = categories
#                         ) , 
#                     by = 'categoryCombo.id'
#                     ) %>%
#     
#         # dataElementGroups
#         left_join( 
#             md$dataElementGroups %>% 
#                 select( name, dataElements) %>%
#                 rename( Group = name ) %>% 
#                 unnest ,
#             by = 'id'
#             )
#     
#     
# 
#     kable( select(mal.conf , name, Group, zeroIsSignificant, lastUpdated, categoryCombo ), 
#            'html' ,
#            caption = "Confirmed malaria data elements" )  %>%
#         kable_styling(bootstrap_options = c("striped", "hover")) %>%
#         column_spec(1, bold = T) 
#     

 # # suspected malaria elements  ####       
 #    mal.susp = de[ mal & susp, display.vars ] %>%
 #        rename( categoryCombo.id = categoryCombo ) %>%
 #        mutate( lastUpdated = format( ymd_hms( lastUpdated ), '%b %Y' ) ) %>%
 #        left_join(  
 #            select( md$categoryCombos, id, name, categories ) %>%
 #                        rename(categoryCombo.id = id,
 #                               categoryCombo = name ,
 #                               category.id = categories
 #                        ) , 
 #                    by = 'categoryCombo.id'
 #                    ) 
 #    
 #    kable( select( mal.susp, name, zeroIsSignificant, lastUpdated, categoryCombo ), 
 #           'html' , 
 #           caption = 'Suspected malaria data elements' ) %>%
 #        kable_styling(bootstrap_options = c("striped", "hover")) %>%
 #        column_spec(1, bold = T) 

```


NB: more description of total vs details for each data element--need to list 

## Datasets


```{r datasets }

  # All dataElements and categoryCombos associated with each dataset
  de = md$dataSets %>% 
    select( id, name, periodType ) %>%
    mutate( categoryCombo = map( md$dataSets$dataSetElements ,
                                 'categoryCombo' ) ,
            dataElement = map( md$dataSets$dataSetElements , 
                               'dataElement' ) 
            ) 

   find_dataset_for_element = function( element ){
       de %>%
        mutate( 
            a = map_lgl( dataElement, ~ any( element %in%  .x$id  ) )
            ) %>%
        filter( a ) %>% select( id )
   } 
   
   # test
   # find_dataset_for_element( mal.conf[1, 'id' ] )

   
   # NB : There may be duplicate rows in de.mal because the same elements belong to multiple groups.
   # Compare with de.mal %>% select(-Group) %>% distinct()
   
    
   mal.datasets.de = de.mal %>% 
       select( id, name ) %>%
       distinct() %>%
       rename(
           dataElement.id = id ,
           dataElement.name = name
       ) %>% 
       # data_frame( 
       #     dataElement.id =  de.mal[, 'id' ] ,
       #     dataElement.name =  de.mal[, 'name' ] 
       #         ) %>%
       mutate( 
           dataset.id = map(
               dataElement.id,  ~find_dataset_for_element( .x ) ) 
           ) %>%
       unnest %>%
       left_join( select(md$dataSets, id, name, periodType ), 
                  by = c('id' = 'id') ) %>%
       rename( dataset.id = id , dataset.name = name ) %>%
       arrange( dataElement.name, dataset.name)
   
   
   # just list of malaria datasets
   mal.datasets = mal.datasets.de %>%
       group_by( dataset.id , periodType ) %>%
       summarise( Name = unique( dataset.name ))%>%
       rename( id = dataset.id, Frequency = periodType ) %>%
       select( Name, id, Frequency) %>%
       arrange( Name )
   
   
   kable( mal.datasets.de, 'html' ) %>% 
       kable_styling(bootstrap_options = c("striped", "hover"))%>%
        column_spec(1, bold = T) 
   
   # NB TODO:  Table of data groups (because it is removed from mal.datasets, to avoid duplicates)

```


 
```{r}

  dataset.ous = md$dataSets %>% 
    select( id, name ) %>%
    # filter( name %in% unique( mal.datasets$dataset.name ) ) %>%
    mutate(
        ou = map( id, ~md$dataSets[  md$dataSets$id==.x , ]$organisationUnits ) ,
        num.ou = map_int( ou , ~length( flatten( .x[[1]] ) ) )
    ) %>%
    select( -ou )

    
    

```
 
Count the org units assigned to each form, which is the basis for the number of expected reports.  If an org unit is not assigned to the dataset, a report is not expected. 

```{r dataset_ous , results = 'asis' }

# function to create html table of a dataset's features
  dataset.ous.n = function( dataset ){
  
    opd.orgs = md$dataSets[ md$dataSets$name %in% dataset ,
                           c( 'name', "organisationUnits" ) ] %>% 
      rename( dataset = name ) %>%
      unnest %>% 
      right_join( select( ous , id, name, level ) , by = "id" )  %>%
      rename( ous = name ) %>%
      # level names
      left_join( 
          select( md$organisationUnitLevels, level, name  ), 
          by = 'level' ) %>%
      rename( levelName = name )
    

    t = count( opd.orgs, level, levelName, dataset ) %>% 
      spread( dataset, n ) %>%
      rename( Unassigned = `<NA>` ) %>%
      kable( "html", caption = dataset ) %>%
      kable_styling(bootstrap_options = c("striped", "hover"))%>%
        column_spec(1, bold = T) 
    
    return(t)
  }

# Create table for each dataset
 for ( i in 1:nrow( mal.datasets ) ){
   cat( dataset.ous.n( mal.datasets[i, 'Name'] ) )
 } 
 
  
```

<!-- - Dataset ous by county.... -->


# Data, Data completeness, reporting trends 



<!-- NB - After getting data, run algorithm for major outliers.  For subsequent charts looking at trends, etc, Replace the outliers so that scale of charts is not distorted to the point where all the true data is not compressed to the bottom of the chart.   -->

```{r fetch_data_values }

  data.file = paste0( folder , 
                     params$origin_login_file, 
                     "_dataset.rds" )

    if ( file.exists( data.file ) ){
        
          data = readRDS( data.file )
          
    } else {
  
    # data elements:
    de.mal.total = paste( de.mal$id , collapse  = ";" )

      # login in to server
      retry( login() )
      
      # import month by month
      
      data = list()
      data.level = list()
      for ( i in seq_along(periods.vector) ){
    
          print( paste( periods.vector[i] ) )
        
        for ( level in seq_along( levels.vector ) ){
          
          print( levels.vector[level] )
      
        #Assemble the URL ( before starting, double check semicolons for dx dimension )
        url <- paste0( baseurl, "api/analytics/dataValueSet.json?" ,
                      "&dimension=ou:", levels.vector[level] , 
                      "&dimension=pe:" , periods.vector[i] ,
                      "&dimension=dx:" , 

                      # malaria
                      de.mal.total ,
                      # opd summary
                      # "KNrK5VWTZkx;NLKRV7bYbVy" , # population
                      "&displayProperty=NAME")
      
        # Fetch data
        fetch <- retry( get(url)[[1]] ) # if time-out or other error, will retry 
        
        # if returns a data frame of values (e.g. not 'server error'), then keep
        if ( is.data.frame( fetch ) )  data.level[[level]] = fetch
        
        } 
        
    # combine level data
    data[[i]] = data.table::rbindlist( data.level )
    
        print( paste( scales::comma( nrow( data[[i]] ) ) , "records"  ) )
    
    }
  
      # combine period data
      data = data.table::rbindlist( data )
      
      print( paste( "TOTAL:", scales::comma( nrow( data ) ) , "records"  ) )
    
      # glimpse(data)
      
      saveRDS( data, data.file )
}

```


## Dataset completeness 

Values shown are for the periods from `r min(periods.vector)` to `r max(periods.vector)`.

```{r fetch_dataset_completeness  }

  data.file = paste0( folder , 
                     params$origin_login_file, 
                     "_datset_completeness.rds" )

    # open if already exists 
    if ( file.exists( data.file ) ){
        
          dataset = readRDS( data.file )  %>% as.tibble 
          
    } else {
  
    # datasets
    datasets =  unique( mal.datasets$id )
    
    reports = c( 'ACTUAL_REPORTS', 'ACTUAL_REPORTS_ON_TIME', 'EXPECTED_REPORTS' )
      
      # login in to server
      login()
      
      # import month by month
      
      dataset.data = list()
      dataset.level = list()
      for ( i in seq_along( datasets ) ){
    
          print( paste( datasets[i] ) )
        
        for ( level in seq_along( levels.vector ) ){
          
          print( levels.vector[level] )
      
        #Assemble the URL ( before starting, double check semicolons for dx dimension )
        url <- paste0( baseurl, "api/analytics/dataValueSet.json?" ,
                      "&dimension=ou:", levels.vector[level] , 
                      "&dimension=pe:" , paste( periods.vector, collapse = ';') ,
                      "&dimension=dx:" , 
                      paste0( datasets[i], ".", reports , collapse = ';')
                       ,
                      "&displayProperty=NAME")
      
        # Fetch data
        fetch <- retry( get(url)[[1]] ) # if time-out or other error, will retry 
        
        # if returns a data frame of values (e.g. not 'server error'), then keep
        if ( is.data.frame( fetch ) )  dataset.level[[level]] = fetch
        
        } 
        
    # combine level data
    dataset.data[[i]] = data.table::rbindlist( dataset.level )
    
        print( paste( scales::comma( nrow( dataset.data[[i]] ) ) , "records"  ) )
    
    }
  
      # combine period data
      dataset = data.table::rbindlist( dataset.data ) %>% as.tibble
      
      print( paste( "TOTAL:", scales::comma( nrow( dataset) ) , "records"  ) )
    
      # glimpse(data)
      
      saveRDS( dataset, data.file )
    }

    d = dataset %>% 
        separate( col = dataElement , into = c('dataSet', 'Category' ), sep = "\\.") %>%
        inner_join( mal.datasets , by = c( 'dataSet' = 'id') ) %>% 
        count( Name, Category ) %>%
        spread( Category, n )
    d
        

```

- Dataset 'completeness' (better described as submission)

```{r data_submission}

    s = dataset %>%
        separate( dataElement, 
                  c('dataElement.id', 'Category' ), sep = '\\.' 
                  ) %>%
        left_join(  select( ous , id, level) , 
                    by = c('orgUnit' = 'id') ) %>%
        group_by( level, period , dataElement.id , Category ) %>%
        summarise( value = sum( as.integer( value) ) ) %>%
        ungroup() %>% 
        spread( Category, value ) %>%
        mutate( 
            submitted = ACTUAL_REPORTS / EXPECTED_REPORTS ,
            onTime = ACTUAL_REPORTS_ON_TIME / EXPECTED_REPORTS 
            ) %>%
         # level names
         left_join( 
             select( md$organisationUnitLevels, level, name  ), 
             by = 'level' 
             ) %>%
        mutate( 
            name = fct_reorder( name, level )
        ) %>%
        rename( level.name = name ) %>%
    
        # data element names 
        left_join( 
            select( md$dataSets, id, name ) %>% 
            rename( dataElement.id = id , dataset = name ) ,
            by = 'dataElement.id' 
            )

    # convert month-Yr to date
    s$date = fast_strptime( as.character(  s$period ) , "%Y%m") 
  
    # summary( s$submitted )
  

```

```{r submission_charts}

# expected number
    ggplot(s,  aes(x = date, y = EXPECTED_REPORTS, 
                   group = level.name , color = level.name )) +
    geom_line() +
    facet_wrap( ~ dataset, 
                labeller = label_wrap_gen(width = 25, multi_line = TRUE)
                ) +
    labs( title = 'Expected reports by level', 
             subtitle =  '' ,
             caption = params$origin_login_file
             )

# submission rate
    ggplot(s,  aes(x = date, y = submitted, 
                   group = level.name , color = level.name )) +
    geom_line() +
    facet_wrap( ~ dataset, 
                labeller = label_wrap_gen(width = 25, multi_line = TRUE)
                ) +
    labs( title = 'Dataset submission rate by level', 
             subtitle =  'denominator is expected reports' ,
             caption = params$origin_login_file
             )
    
# onTime rate
    ggplot(s,  aes(x = date, y = onTime, 
                   group = level.name , color = level.name )) +
    geom_line() +
    facet_wrap( ~ dataset, 
                labeller = label_wrap_gen(width = 25, multi_line = TRUE)
                ) +
    labs( title = 'Dataset submission-on-time rate, by level', 
             subtitle =  'denominator is expected reports' ,
             caption = params$origin_login_file
             )
    
```

link with org units 

```{r read_dataset_completeness }

  ous_complete = md$organisationUnits %>% 
    select( id, name, path , featureType, 
            # active , # is not available in all API ?
            openingDate, closedDate ) %>%
    mutate( level = map_dbl( path, 
                           ~length( gregexpr("/", .x, perl = TRUE)[[1]]) 
                           ) 
    )
  
 
```

## Completeness

Completeness by level: 

- crude measure with denom = all orgunits

link with org units 

```{r read_data }

  ous = md$organisationUnits %>% 
    select( id, name, path , featureType, 
            # active , # is not available in all API ?
            openingDate, closedDate ) %>%
    mutate( level = map_dbl( path, 
                           ~length( gregexpr("/", .x, perl = TRUE)[[1]]) 
                           ) 
    )
  
 
```

```{r completeness_crude }
  
  d.ous = inner_join( data, ous, by = c( 'orgUnit' = 'id' ) )
                     
  d.ous.noCat = d.ous %>% 
      group_by( period, level , orgUnit, dataElement ) %>%
      summarise( value = sum( as.integer(value), na.rm = TRUE ) ) %>%
      ungroup %>%
      # level names
      left_join( 
          select( md$organisationUnitLevels, level, name  ), by = 'level' )
    
  s = d.ous.noCat  %>%
    mutate( 
        name = fct_reorder( name, level )
        ) %>%
    rename( dataElement.id = dataElement , 
            level.name = name ) %>%
    group_by( period, level, level.name , dataElement.id ) %>%
    summarise( report = n() )  %>% 
    left_join( count( ous, level ) , by = 'level' ) %>%
    mutate( completeness = report / n ) %>%
    ungroup() %>%
    # data element names 
    left_join( 
        select( md$dataElements, id, name ) %>% 
            rename( dataElement.id = id , dataElement = name ) ,
        by = 'dataElement.id' 
    )
  
  # convert month-Yr to date
  s$date = fast_strptime( as.character(  s$period ) , "%Y%m") 
  
  summary( s$completeness )
  
    ggplot(s,  aes(x = date, y = completeness, 
                   group = level.name , color = level.name )) +
    geom_line() +
    facet_wrap( ~ dataElement, 
                labeller = label_wrap_gen(width = 35, multi_line = TRUE)
                ) +
    labs( title = 'Crude completeness by data element', 
             subtitle =  'denominator is current number of all org units' ,
             caption = params$origin_login_file
             )
  
```

- Summarise org units open each period

Limitation: Because only one openiningDate is recorded, if clinic was closed, then reoppened, it will be reported as if it was always open.  


```{r open_orgs}
 
 
    # convert month-Yr to date
  d = ous.n %>%
      # level names
      left_join( 
          select( md$organisationUnitLevels, level, name  ) , 
          by = 'level' ) %>%
    mutate( 
        name = fct_reorder( name, level )
        )

  d$date = fast_strptime( as.character(  d$period ) , "%Y%m") 
  
  ggplot( d,  aes(x = date, y = n, group = level )) + 
      geom_line() +
      facet_wrap( ~name , scales = 'free' )
  

  
```

<!-- NB:  Use API to save charts as favorites??? -->

<!-- Compare with indicators from other programs--tb, hiv, imm... -->

## Map: Health facilities with number of confirmed malaria cases 


```{r}
# orgUnits that are points (likely HF), and open
clinics = md$organisationUnits %>% 
    select( id, name, coordinates, openingDate, closedDate, path  ) %>%
    as.tibble %>% 
    mutate( 
        feature_type = map_chr( coordinates, ~feature_type(.x) ) ,
        # remove quotation marks
        coordinates = gsub('"', '', coordinates)
        ) %>%
    filter( feature_type %in% 'Point') %>%
    separate( coordinates, c('long', 'lat'), sep = ',', remove = FALSE ) %>% 
    filter( !is.na(long), !is.na(lat), is.na( closedDate ) ) %>%
    mutate(
        long = as.numeric(  substr(  long, 2, nchar(long) ) ) ,
        lat =  as.numeric(  substr(  lat, 1, nchar(lat)-1 )  ) ,
        level = map_int( path,
                         ~length( gregexpr("/", .x, perl = TRUE)[[1]]) 
            )
    ) 

# NB TODO: Give message for coordinates missing lat long
```


```{r}
# 2017 data
    data.2017 = data %>%
        filter( orgUnit %in% clinics$id ) %>% 
        mutate( year = substr( period , 1, 4 )  )  %>%
        filter( year %in% 2017 ) %>%
        group_by( year ,  orgUnit, dataElement ) %>%
        summarise( value = sum( as.double(value) ) ) %>%
        left_join( select( md$organisationUnits , id , name  , path) %>%
                       mutate( level = map_int( path,
                                                ~length( gregexpr("/", .x, perl = TRUE)[[1]]) 
                       ) ),
                   by = c("orgUnit" = 'id' )
                   ) %>%
        ungroup %>%
        rename( HF = name, id = dataElement ) %>%
        left_join( select( md$dataElements, id, name ) ,
                   by = c( 'id' ))  %>%
        rename( dataElement = name ) %>%
        select( year , level, HF, dataElement, value ) 
    
# numbers of HF submitting data (note, does not reflect number of months.  could be 1 or 12)
    # count(data.2017, level, dataElement)
```

```{r, fig.width= 8, fig.height=10 }

# select data.element to show by finding most frequent
de_freq = data  %>%
    mutate( year = substr( period, 1, 4 )) %>%
    inner_join( select( md$dataElements, id, name  ) , 
                by = c('dataElement'='id' )) %>%
    count( year, name ) %>%
    arrange( desc(year), desc(n) ) 

de_most_freq = de_freq[1, ]


d = left_join( clinics , 
                  filter( data.2017, 
                          dataElement %in% de_most_freq 
                          # value > 0 
                          ) %>%
                      select( year, HF, value ),  
                  by = c('name' = 'HF') 
                              ) %>%
                mutate( 
                    value = ifelse( is.na( value ), 0 , value ) ,
                    value.scale = ifelse( !is.na(value) & value>0 , 
                                                    1 + log(1 + value/ mean(value, na.rm = TRUE) ) , 
                                                 2 ) 
                        ) %>%
    filter( abs(long) <= 180 , abs(lat) <= 180 )

 # NB: add test to see if coords within country; if not, try reversing


    pal = colorNumeric('OrRd', d$value.scale )
    binpal <- colorBin("plasma", d$value , 5, pretty = TRUE)
    factpal <- colorFactor( brewer_pal("qual", 2)(7),  d$level)

 leaflet( d ) %>% 
        addTiles(  urlTemplate = 
                       "http://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png"
                   )  %>%
        addCircles( 
            ~long, ~lat, 
            radius =  500 ,
            weight =  ~value.scale ,
            label = ~paste( name , "(level:", level, ")" , comma( value ) )   ,
            labelOptions = labelOptions(noHide = FALSE, textOnly = TRUE, textsize = "14px") ,
            popup = ~paste( name , "(level:", level, ")" , comma( value ) ) , 
            color = ~binpal(value)  , opacity = 1 ,
            fillColor = ~binpal(value) , fillOpacity = .6
        ) %>%
        addScaleBar() %>%
        # addLegend(pal = factpal, values = ~level , opacity = 1 ) %>%
        addLegend(pal = binpal, values = ~value, opacity = 1 ) 
 

```


## Submissions by data element / year

Percentages shown in each chart are  
1. Percentage of data elements with non-missing value   
2. Percentage of OU that submitted 12/12 monthly reports


```{r submisison_by_dataElement}

ou_expected =  dataset %>% as.tibble %>%
        separate( dataElement, c('dataSet', 'Category' ), 
                  sep = '\\.' 
                  ) %>%
        filter( Category %in% 'EXPECTED_REPORTS', 
                as.integer(value) == 1L
                ) %>% 
        # levels
        left_join( ous %>% rename( orgUnit.name = name ), 
                   by = c('orgUnit' = 'id')  ) %>%
        # level names
         left_join( 
             select( md$organisationUnitLevels, level, name  ) %>%
                 rename( level.name = name ), 
             by = 'level' 
             ) %>%
    select( dataSet, Category, period, orgUnit, orgUnit.name , level.name )
        
# add expected reported to data, call new dataset 'submission'
submission = data %>% as.tibble %>% 
    mutate( year = substr( period, 1, 4 )) %>%
    # dataset names
    left_join(
        mal.datasets.de  , 
        by = c('dataElement' = 'dataElement.id' )
        # NB: may have multiple datasets for one dataElement
    ) %>% 
    # link with facilities expected to report
    inner_join( 
        ou_expected , 
        by = c( 'period'='period',  
                'dataset.id' = 'dataSet', 
                'orgUnit' = 'orgUnit') 
        )

submission_annual = submission %>% 
    group_by( 
        level.name, dataset.name, dataElement, dataElement.name, year, orgUnit 
        ) %>%  
    summarise( submission = sum( !is.na( value )) ) %>%
    ungroup %>%
    mutate(
        dataset_dataElement = paste( dataset.name, dataElement.name,  sep = '_' )
    )

# NB: TODO FIx these summaries so they works for partial years

submission_annual_rate = submission_annual %>%
    group_by( dataset.name, dataElement.name, dataset_dataElement, year ) %>%
    summarise( submission_rate = mean( submission / 12 ) ,
               percent_12 =  sum(submission == 12) / n() ) %>%
    ungroup()

# submission_annual_rate %>% 
#     select( dataElement.name, dataset.name, year, submission_rate ) %>%
#     arrange( desc(year), desc(submission_rate) ) 

most_submitted_dataset = submission_annual_rate$dataset.name[1]


```

<!-- The histograms below are for data elements associated with **`r most_submitted_dataset`**, the most frequently submitted dataset. -->

```{r submission_histograms, fig.width=10, fig.height=8 }

d =  count( submission_annual ,
            # filter( submission_annual , 
            #     dataset.name %in% most_submitted_dataset 
            #     ) ,
        dataset_dataElement , year, submission ) 

ggplot( d ) +
    geom_bar( aes( factor( submission ) , n), stat = 'identity' ) + 
    geom_text( data = submission_annual_rate ,
               # %>% 
               #     filter( dataset.name %in% most_submitted_dataset )  , 
               aes(label = paste(
                       percent(submission_rate) , 
                       percent( percent_12 ) ,
                       sep = "\n" 
                       )
                   ) ,  x= 2, y= .8 * max( d$n, na.rm = TRUE ), 
               hjust=0, size = 3 ) +
    labs( x = "Number of submisisons per year", y = "Number of facilities" ,
          title = 'Histogram of the number of monthly submissions per year' ,
          subtitle = '(range 0-12 / yr)' ) +
    facet_grid(  year ~ dataset_dataElement  ,
                 labeller = label_wrap_gen(width = 20, multi_line = TRUE) )

```

<!-- TODO: Map submission rate for the previous year (12 months?) by district -->

## Submission consistency

Number of facilities reporting every month for the last 48 months, by dataset and data element. 

```{r continuous_submisison}

# function to determine which facilities reporting during previox xx intervals

continuous = function( data = submission , intervals = 24 ){
    
    periods = unique( data$period )
    periods = periods[ order(periods) ]
    n_periods = length( periods )
    
    if (n_periods < intervals){
        print( paste(
            'There are only', n_periods, 'intervals; the function will test for contiunous submission for up to', n_periods, 'intervals.' 
            ) )
        
        intervals = n_periods
        
    }
    
    interval_period = periods[ (n_periods-intervals + 1):n_periods ]
    
    s = data %>%
        # de name
        left_join(
            select( md$dataElements , id, name ) ,
            by = c('dataElement' = 'id' )
            ) %>%
        filter( 
            period %in% interval_period
                ) %>% 
        group_by(
            dataElement.name , dataset.name, orgUnit 
            ) %>%
        summarise( 
            n = n()
            ) %>%
        mutate( 
            all = factor( n == intervals ,
                          levels = c( TRUE, FALSE ) ,
                          labels = c("Clinics continuosly reporting" , 
                    'Other clinics' )
                    )
        )
        
    
    return(s)
}

 d = submission  %>% 
    continuous( intervals = 24 ) %>%  
    left_join(
        select( ous , id, level ) , 
        by = c('orgUnit' = 'id')
    ) %>% 
    group_by( level, dataElement.name , dataset.name ) %>%
    summarise( all = sum( as.numeric(all)) )  %>%
    spread( dataset.name, all ) 
 
 kable( d, 'html' ) %>% 
       kable_styling(bootstrap_options = c("striped", "hover"))%>%
        column_spec(1, bold = T) 

```

```{r, eval = FALSE}

map( c(12, 24, 36, 48), ~continuous( submission , intervals = .x) %>%
         summarise( all = sum(all) ) ) %>% View

```

## Trends

Compare trends among clinics reporting consistently for the last **24 months** with all the other clinics, many of which started reporting more recently. 

Left column is the number of reported cases

Right column is the number of facilities reporting per month

```{r}

# scale function
scale_value <- function(x){
  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
}

trend = submission %>% 
    unite( data.dataSet , dataElement.name, dataset.name, 
           remove = FALSE ) %>%
    # filter( 
    #     # grepl( 'opd confirmed' , dataElement.name, ignore.case = TRUE ),
    #     grepl( 'malaria' , dataset.name, ignore.case = TRUE )
    #     ) %>% 
    left_join( continuous( intervals = 24 )  ,
               by = c( 'orgUnit', 'dataElement.name', 'dataset.name' )
               ) %>%
    group_by( data.dataSet, period, all ) %>%
    summarise(
        value = sum( as.integer( value ) ) , 
        n = n() 
    ) %>% 
    group_by( data.dataSet, all ) %>%
    mutate( 
            value_rolling = rollmean( value  , 12, 
                                       align = 'right', na.pad = TRUE) ,
            value.scale = scale_value( value ) ,
            value_rolling.scale = scale_value( value_rolling ) 
        )     %>% 
    ungroup

trend$date = fast_strptime( as.character( trend$period ) , "%Y%m") 

# number of variables...
num_data.dataSet = length( unique( trend$data.dataSet ) ) 

```


```{r trend_plots, fig.width = 8 , fig.height = 1.5 * num_data.dataSet }

# ggplot
trend.value =  ggplot( data = trend  ) +
    geom_line( aes( x = date, y = value , 
                    group = all, color = all )
               ) +
    theme(legend.position="bottom") +
    labs( y = 'Cases', title = "Number malaria cases" , subtitle= "by dataElement_dataSet" ) +
    facet_wrap( ~data.dataSet, scales = 'free' , ncol = 1, 
                 labeller = label_wrap_gen(width = 50, multi_line = TRUE)
                )

trend.n =  ggplot( data = trend  ) +
    geom_line( aes( x = date, y = n , 
                    group = all, color = all )
               ) +
    theme(legend.position="bottom") +
    labs( y = 'Number submissions', title = "Clinics submitting malaria cases" , subtitle= "by dataElement_dataSet" ) +
    facet_wrap( ~data.dataSet, scales = 'free' , ncol = 1, 
                 labeller = label_wrap_gen(width = 50, multi_line = TRUE)
                )

grid.arrange( trend.value , trend.n,  ncol=2 )

```


## Signal to noise ratio 

<!-- https://www.socialresearchmethods.net/kb/expclass.php -->

<!-- http://www.stat.columbia.edu/~gelman/stuff_for_blog/chap20.pdf -->

<!-- 'Observable effect size'  comparing annual cases by region and significant trends -->

# Field visits

- site selection options
