---
output:
  html_document:
    # css: custom.css
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
params:
  dhis_instance: WHO Demo
  data_directory:  .
  output_directory: ./ # bug: only uses current directory
  cache: FALSE
  echo: FALSE
---

# **`r params$dhis_instance` Trends** {.tabset .tabset-fade}
<!-- load libraries... -->

## Outline {.tabset .tabset-fade}

- Identify most important dataElements and dataSets (do this nationally, and by county)

1. Find HF that have been continuously reporting; compare the trend with rest of the data   
    a. sum case counts by month   
    b. Use seasonal decomposition models to estimate past trends 
    c. Forecast
    
2. Calculate mean case count by month across all clinics (assumes that clinic case counts are normally distributed)
    
3. Calculate incidence with available population data--if it is by health facility

4. Calculate case rate per 1,000 patients attending each HF, then take mean of across all HF

5. Impute missing values based on location, # patients


```{r packages, message=FALSE, include= FALSE }

# list of required packages
    package.list = c( "openxlsx", "readxl" , "rlist", "knitr", "kableExtra", "xts", "leaflet", "RColorBrewer", "DT", "dygraphs", "httr", "jsonlite", "XML", "assertthat", "lubridate", "scales", "RcppRoll", "zoo", "gridExtra", "futile.logger", "zoo" , "broom" , "Hmisc", "skimr",  "tidyverse" )

# Function to test if package is installed 
    pkgTest <- function( package.list = package.list ){
        
        missing.packages = setdiff( package.list , rownames(installed.packages())) 
        if ( length( missing.packages ) > 0 ) install.packages( missing.packages ) 
    }


# Test if packages loaded
    pkgTest( package.list )

# load the packages
    lapply( package.list, suppressMessages( require ) , character.only = TRUE)


knitr::opts_chunk$set( echo = params$echo ,
                       fig.width =  8, 
                       # knitr.table.format = "html" ,
                       cache = params$cache # TRUE for testing; FALSE for production
                       )

```

```{r sources }

source('dhis2_functions.R')

    
    data_directory = params$data_directory
    dhis_instance = params$dhis_instance
    output_directory = params$dhis_instance
  
  # folder for storing review data

  origin.folder = paste0( ifelse( is.na( data_directory ), "" , 
                                  paste0( data_directory, "/" ) 
                                  ) ,
                          dhis_instance , "/")
  
 # login shortcuts

login_origin = function(){
    source( 
    paste0( origin.folder, 
            tolower( params$dhis_instance) , 
            "_login" )
    )

  loginDHIS2( baseurl, username, password)
}
  
```

```{r all_meta }

  meta_data_file = paste0( origin.folder ,
                           dhis_instance, 
                           "_metadata.rds" )

  if ( !file.exists( meta_data_file) ){
   
    # login in to server
    login_origin()
  
    md = metadataDHIS2( baseurl, 'all' )
    
    # glimpse( md )
    
    write_rds( md, meta_data_file )
    
  } else {
      
       md = read_rds( meta_data_file )
  }

  # re-order alaphabetically to be easier to browse in View
  md = md[order(names(md))]
  
  # Retrieve access data 
  date_metadata = file.info( meta_data_file )$ctime 
  
```

```{r date_strings }

 # Function to create string of dates
  date_code = function( years = 2013:2017 , months = 1:12 ){
    period = character()
    
    for (year in years ){
      for (month in seq_along(months)){
        this_period = paste0( ";", 
                         year , 
                         ifelse( month < 10 , paste0("0", month) , month )
        )
        period =  c(period, this_period )
      }
    }
    
    # remove first ;
    period =  paste( period, collapse = "")
    period = substring( period, 2, nchar(period))
    
    return( period )
  }
 
  periods = date_code() # default: 2013-2017
  # periods = date_code( 2017, 1 ) # January 2017 only
  
  # as character vector
  periods.vector = strsplit( periods, ";" , fixed = TRUE )[[1]]

  
```

```{r ous_metatdata }

    feature_type = function( coordinate ){
        n = length( gregexpr( '[' , coordinate , fixed = TRUE)[[1]] )
        
        if ( is.na( coordinate ) ) return(NA)
        if (n==1) return('Point') 
        if (n>1) return('Polygon') 
        
    
    }

  parse_parent_ous = function( path ){
       breaks = gregexpr("/", path , perl = TRUE)[[1]]
       n = length( breaks ) 
       if ( n == 0 ) return( NA ) 
       if ( n == 1 ) return( substr( path , breaks[1] + 1 , length(path) ) ) 
       substr( path , breaks[n-1] + 1 , breaks[n]-1 )
  }
  
   ous =  md$organisationUnits %>% 
        select( id, path, name, shortName, coordinates , 
                created, lastUpdated, ends_with('Date') 
                ) %>% as.tibble %>%
       rename( ous = name ) %>%
        mutate( 
          ous = ifelse( ous %in% "", shortName, ous ) , #if name blank, use shortName
          level = map_chr( path, 
                               ~length( gregexpr("/", .x, perl = TRUE)[[1]]) 
                               ) ,
          parent.id =  map_chr( path, ~parse_parent_ous( .x ) ) ,
          feature = map_chr( coordinates, ~feature_type(.x ) )
        ) %>%
       # Name of parent
       left_join( md$organisationUnits %>% select( id, name ) ,
                   by = c('parent.id' = 'id' ) ) %>%
       rename( parent = name ) %>%
       # Name of level
       left_join( md$organisationUnitLevels %>% select( level, name ) %>%
                      mutate( level = as.character( level ) ),
                  by = 'level' ) %>%
       rename( level.name = name)
 
``` 

## DataElements : Totals 

```{r datasets  }

    # Totals
    data.totals.file = paste0( origin.folder , 
                     dhis_instance, 
                     "_dataset_totals.rds" )
    data_totals = readRDS( data.totals.file ) 
    data_totals_date = file.info(data.totals.file)$mtime
    
    # Details
    data.details.file = paste0( origin.folder , 
                     dhis_instance, 
                     "_dataset_details.rds" )
    data_details = readRDS( data.details.file )
    data_details_date = file.info(data.details.file)$mtime
    
    # Submissions
    data.submissions.file = paste0( origin.folder , 
                     dhis_instance, 
                     "_dataset_submissions.rds" )
    data_submissions = readRDS( data.submissions.file ) 
    data_submissions_date = file.info(data.submissions.file)$mtime
```

- The totals data was accessed on `r data_totals_date`  
- The detailed data was accessed on `r data_details_date`  
- The submissions (dataset) information was downloaded on `r data_submissions_date`  

## Data elements associated with datasets

```{r dsde }

  # data frame of datasets (ds) and data elements (de)
  dsde = map_df( 1:length(md$dataSets$dataSetElements), 
            ~map_df( md$dataSets$dataSetElements[[.x]], 
                     ~as.matrix(.x) ))

    # Data elements in data.totals
    de = data_totals %>% select( dataElement ) %>% distinct 
    
    # List associated datasets with names
    datasetsWithDE = dsde %>%
        inner_join( de, by = 'dataElement' ) %>%
        left_join( md$dataSets %>% select( id, name ), 
                   by = c("dataSet"="id")
                   ) %>%
        select( -dataSet ) %>%
        rename( dataSet = name ) %>% 
        left_join( md$dataElements %>% select( id, name ), 
                   by = c("dataElement"="id")
                   ) %>%
        select( -dataElement ) %>%
        rename( dataElement = name ) %>% 
        left_join( md$categoryCombos %>% select( id, name ), 
                   by = c("categoryCombo"="id")
                   ) %>%
        select( -categoryCombo ) %>%
        rename( categoryCombo = name )
    
    # display 
        datasetsWithDE %>%
        datatable( rownames = FALSE,
                   options= list( pageLength = 15 ,
                                  scrollY = TRUE)
                   ) 
        
        
```

```{r dataset_withDE_list}

 dataSets_to_Download = datasetsWithDE %>% select(dataSet) %>% distinct

  kable( dataSets_to_Download )

```

# Submission consistency {.tabset .tabset-fade}

Number of facilities reporting every month for the last 48 months, by dataset and data element. 

## Numbers of org units

```{r ou_levels}

  # levels
      levels = character()
      for ( i in seq_along( md$organisationUnitLevels$id) ){
      # for ( i in seq_along( count(ous, level) %>% .$level ) ){
    
        levels =  c(levels, paste0("LEVEL-", i) )
      }
      
      levels = paste( levels, collapse = ";")
  
  # import month by month
  levels.vector = strsplit( levels, ";" , fixed = TRUE )[[1]]
  
```


```{r continuous_submisison}

# function to determine which facilities reporting during previox xx intervals

continuous = function( data = submission , months = 24 ){
    
    periods = fast_strptime( as.character( data$period ) , "%Y%m") 
    max_period = max(periods, na.rm = TRUE)
    first_month_in_interval = max_period %m-% months( months )
    
    continuous.interval = interval( first_month_in_interval , max_period ) 

    available.interval = interval( min(periods, na.rm = TRUE) , 
                               max(periods, na.rm = TRUE) ) 
    
    n_periods = time_length( available.interval , 'months')
    
    
    if (n_periods < months){
        print( paste(
            'There are only', n_periods, 'intervals; the function will test for contiunous submission for up to', n_periods, 'intervals.' 
            ) )
        
         continuous.interval = available.interval 
        
    }
    

    s = data %>%
        
        # Remove any rows with all missing data
        select( -starts_with( 'NA' ) ) %>%
        filter( complete.cases(.) ) %>%
            
        mutate( dates = fast_strptime( as.character( period ) , "%Y%m") %>%
                    as.POSIXct()
                ) %>%
        filter( 
            dates %within% continuous.interval
                ) %>% 
        group_by(
            dataElement , orgUnit 
            ) %>%
        summarise( 
            n = n() , 
            continuous = n() >= months
            ) 
    
    
        # mutate( 
        #     all = factor( continuous  ,
        #                   levels = c( TRUE, FALSE ) ,
        #                   labels = c("Clinics continuosly reporting" , 
        #             'Other clinics' )
        #             )
        # )
        
    
    return(s)
}

# number of continuously reporting orgUnits by level for each dataElement 
 d = continuous( data_totals, months = 36 )  %>%  
    left_join(
        select( ous , id, level ) , 
        by = c('orgUnit' = 'id')
    ) %>% 
    left_join( dsde, by = 'dataElement' ) %>%
    group_by( dataSet, dataElement , level ) %>%
    summarise( all = sum( as.numeric( continuous ) ) )   %>%
    ungroup %>%
     
     #dataset name
    rename( id = dataSet ) %>%
    left_join( md$dataSets %>% select( id, name ) ,  by = 'id'  ) %>%
    rename( dataSet = name ) %>%
    select( -id ) %>%
     
    # data element name
    rename( id  = dataElement ) %>%
    left_join( md$dataElements %>% select( id, name ) , by = 'id' ) %>%
    rename( dataElement = name ) %>%
    select( -id ) %>%
     
    spread( level, all ) %>% 
    
     # Sort by most frequent
    rowwise() %>% 
     mutate(
        max = ncol(.) %>% tail(1)
        )  %>%
    arrange( -max ) %>%
    mutate( row = row_number()) %>%
    select( row, dplyr::everything(), -max )
 
 kable( d, 'html' ) %>% 
       kable_styling(bootstrap_options = c("striped", "hover") ) %>%
        column_spec(1, bold = T) 

```

Chart of number of health facilities reporting each data element as function of number of continuous months with data 

```{r chart_continous }

  most_continuous_elements = unique( d$dataElement[1:8] )

  c = map_df( c(1, 6, 12, 24, 36, 48 ), ~continuous( data_totals, .x )  %>%
                     mutate( months = .x ) %>%
                     group_by( dataElement, months ) %>%
                     mutate( n_continuous = sum( continuous ) )
                     
     ) %>%
      
    # add dataset
    left_join( dsde, by = 'dataElement' )  %>%
    
    #dataset name
    rename( id = dataSet ) %>%
    left_join( md$dataSets %>% select( id, name ) ,  by = 'id'  ) %>%
    rename( dataSet = name ) %>%
    select( -id ) %>%
     
    # data element name
    rename( id  = dataElement ) %>%
    left_join( md$dataElements %>% select( id, name ) , by = 'id' ) %>%
    rename( dataElement = name ) %>%
    select( -id ) %>%
      
    # Include only those that are most_continuous_elements
     filter( dataElement %in% most_continuous_elements ) %>%
    
    mutate( de_ds = paste0( dataSet , "__", dataElement ))

    # glimpse(c)
    
    g = ggplot(c, aes( months, n_continuous, group = dataElement , 
                       color = dataElement ) ) +
        geom_point( size = 2 )  +
        geom_line() +
        theme(legend.position='none') +
        facet_wrap( ~dataSet ,
                labeller = label_wrap_gen(width = 25, multi_line = TRUE) 
                ) 
    
    # g
    
    plotly::ggplotly( g )
```


## Trends 

Compare trends among clinics reporting consistently for the last **48 months** with all the other clinics, many of which started reporting more recently. 

This chart show trends at national level (left column) compared with trends when restricting to data from health facilities that reported continously for past 48 months (right column).

```{r }

# scale function
scale_value <- function(x){
  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
}

trend = data_details %>% 
    
    # Include only those orgUnit-dataElement pairs that have continuously reported
    inner_join( continuous( data_totals, months = 48 ) %>%
                    filter( continuous == TRUE ) ,
                by = c( 'dataElement', 'orgUnit' )  ) %>%


    # data element name 
    rename( id =  dataElement ) %>%
    left_join( md$dataElements %>% select( id, name ),
               by = 'id' ) %>%
    rename( dataElement = name ) %>%
    select( -id ) %>%
    
    filter( dataElement %in% most_continuous_elements ) %>%

    # CategoryOptionCombo names
    rename( id =  categoryOptionCombo ) %>%
    left_join( md$categoryOptionCombos %>% select( id, name ) ,
               by = "id" ) %>%
    rename( categoryOptionCombo = name ) %>%
    select( -id ) %>%
    
    unite( "element.combo" , c( "dataElement",  "categoryOptionCombo" ) , 
           sep = "." , remove = FALSE ) %>%
  
    
    # org Unit names
    rename( id =  orgUnit ) %>%
    left_join( ous %>% select( id, ous, level, level.name ) ,
               by = "id" ) %>%
    rename( orgUnit = id ) %>% 
    
    group_by( element.combo, dataElement, categoryOptionCombo, period, level , level.name) %>%
    
    summarise(
        value = sum( as.integer( value ) ) , 
        n = n() 
        
    ) %>% 
    
    group_by( element.combo , dataElement, categoryOptionCombo , level, level.name ) %>%

    mutate( 
            # value_rolling = rollmean( value  , 12,
            #                            align = 'right', na.pad = TRUE)
            # ,
            value.scale = scale_value( value ) 
            # value_rolling.scale = scale_value( value_rolling )
        )     %>% 
    
    ungroup

if (nrow(trend)> 0 ) trend$date = fast_strptime( as.character( trend$period ) , "%Y%m") %>% as.POSIXct()


num_elements = nrow( count( trend, element.combo ))
```


```{r trend_plots, fig.width = 8, fig.height= 12  }

# Value trend at level 1
d = trend %>% filter( level %in% min(.$level)  )

if (nrow(d)> 0 ){ 
    total.trend =  ggplot( data = d ) +
        geom_line( aes( x = date, y = value , group = element.combo
                        )
                   ) +
        theme( legend.position="bottom") +
        labs( y = 'Number', 
              title = "Number malaria cases" , 
              subtitle= "by dataElement_categoryOptionCombo" ) +
        facet_wrap( ~element.combo, scales = 'free' , ncol = 2 , 
                     labeller = label_wrap_gen(width = 20, multi_line = TRUE) 
        ) +
        scale_y_continuous( labels = scales::comma )
    
    # Value trend at lowest reporting level
    d = trend %>% filter( level %in% max(.$level)  )
    
    select.trend =  ggplot( data = d ) +
        geom_line( aes( x = date, y = value , group = element.combo
                        )
                   ) +
        theme( legend.position="bottom") +
        labs( y = 'Number', 
              title = "Number malaria cases" , 
              subtitle= "by dataElement_categoryOptionCombo" ) +
        facet_wrap( ~element.combo, scales = 'free' , ncol = 2 , 
                     labeller = label_wrap_gen(width = 20, multi_line = TRUE)
                    ) +
        scale_y_continuous( labels = scales::comma )
    
    
    grid.arrange( total.trend , select.trend,  ncol=2 )

} else { 
    cat("No data to evaluate")
    }
```


# Signal to noise ratio 

This section will highlight some mechanism for estimating the strength of the surveillance 'signal' relative to the 'noise' inherent in routine health facility data collection.  

<!-- https://www.socialresearchmethods.net/kb/expclass.php -->

<!-- http://www.stat.columbia.edu/~gelman/stuff_for_blog/chap20.pdf -->

<!-- 'Observable effect size'  comparing annual cases by region and significant trends -->
