---
title: "DHIS_Review"
date: "April 6, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  dhis_instance: Kenya
  data_directory:  
  output_directory: ./ # bug: only uses current directory
  output_format: html_document
  create_new: FALSE 
  cache: FALSE
  echo: FALSE
---


```{r package_list, message=FALSE, include= FALSE, warning= FALSE }

# list of required packages
    package.list = c( "openxlsx", "readxl" , "rlist", "knitr", "rmarkdown", "kableExtra", "xts", "leaflet", "RColorBrewer", "DT", "dygraphs", "httr", "jsonlite", "XML", "assertthat", "lubridate", "scales", "RcppRoll", "zoo", "gridExtra", "futile.logger", "zoo" , "broom" , "Hmisc", "plotly" , "skimr", "tidyverse" )

# Function to test if package is installed 
    pkgTest <- function( package.list = package.list ){
        
        missing.packages = setdiff( package.list , rownames(installed.packages())) 
        if ( length( missing.packages ) > 0 ) install.packages( missing.packages ) 
    }


# Test if packages loaded
    pkgTest( package.list )

# load the packages
    lapply( package.list, suppressMessages( require ) , character.only = TRUE)


# load accessory functions
source('dhis2_functions.R')

```


```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set( echo = params$echo ,
                       comment = "" ,
                       cache = params$cache , 
                       fig.width =  8
                       )


    dhis_instance = params$dhis_instance
    data_directory = params$data_directory  # folder containing login_file 
    output_format = params$output_format
    
    # this folder will hold output of the metadata and data reviews
    output_directory =  getwd() # "." is the directory where this file sits. 
    
    origin.folder = paste0( ifelse( is.na( data_directory ), "" , 
                                  paste0( dhis_instance, "/" ) 
                                  ) ,
                          dhis_instance , "/")
    
    if (!dir.exists( origin.folder ) ) dir.create( origin.folder )
    
  
 # login shortcuts

login_origin = function(){
    source( 
    paste0( origin.folder, 
            tolower( params$dhis_instance) , 
            "_login" )
    )

  loginDHIS2( baseurl, username, password)
}

# test login 
 assert_that( login_origin() )
  
    
```


```{r parameters2 }
    params.list = list(
        dhis_instance = dhis_instance , # selects dhis2 instance
        data_directory = data_directory ,  # root data directory
        output_directory = output_directory ,
        cache = FALSE ,
        echo = FALSE 
    )

    outputs = data_frame( 
        output_format = c(  "html_document", 'word_document'),
        output_suffix = c(  "html", 'doc')
    )
    
    output.suffix = outputs %>% 
        filter( output_format == params$output_format ) %>% 
        .$output_suffix
```

# **`r toupper( params$dhis_instance )` **

This script creates the review documents 

## Metadata

```{r Metadata_Review }
    output.filename = paste0( output_directory, "/" ,
                              dhis_instance, "/", 
                             dhis_instance, "_Metadata_Review." ,
                             output.suffix
                             ) 
    if ( !file.exists( output.filename ) | params$create_new ) { 

        render('Review_Metadata.Rmd',
           output_file = output.filename ,
           output_format = output_format ,
           # 'knitrBootstrap::bootstrap_document' ,
           params = params.list ,
           envir = new.env() ,
           encoding = "UTF-8"
           )
         
    }

   if ( file.exists( output.filename )) browseURL( output.filename ) 

```


## Select data elements to review

```{r dataElement_Selection }
    de.review.xlsx = paste0( output_directory , "/" ,
                             dhis_instance , "/" ,
                             dhis_instance , "_" , 
                             "dataElements.xlsx" )
    
    # if exist, open excel file
    if ( file.exists( de.review.xlsx ) ){
        
        # de.include = readxl::read_xlsx( de.review.xlsx  ) %>% 
        # filter( !is.na(Include) ) %>% 
        # distinct( dataElement, .keep_all = TRUE )
        # 
        # if ( nrow( de.include) > 0 ) browseURL( de.review.xlsx )
        
        browseURL( de.review.xlsx )
    }
    
```


<!-- May want to re-select after viewing totals, and before downloading details with categoryCombos -->

## Data Element Review
    
    Data elements selected from the spreadsheet `r de.review.xlsx` will be accessed for review.  If no elements have been selected from the spreadsheet, the revieww will default to data elements that match the search criteria 'malaria' and 'confirmed'. 
    
```{r Data_Review}
    # if variables selected, download and analyze

    output.filename = paste0(output_directory , "/" ,
                             dhis_instance, "/", 
                             dhis_instance, "_Data_Review.html" ) 

     if ( !file.exists( output.filename ) | params$create_new ) { 

        render('Review_Data.Rmd',
           output_file = output.filename ,
           output_format = output_format ,
           # 'knitrBootstrap::bootstrap_document' ,
           params = params.list ,
           envir = new.env() ,
           encoding = "UTF-8"
           )
         
    }

   if ( file.exists( output.filename )) browseURL( output.filename ) 


```

## Trends and Quality
    
```{r Trend_Review}
    # if variables selected, download and analyze

    output.filename = paste0(output_directory , "/" ,
                             dhis_instance, "/", 
                             dhis_instance, "_Trend_Review.html" ) 


     if ( !file.exists( output.filename ) | params$create_new ) { 

        render('Review_Trends.Rmd',
           output_file = output.filename ,
           output_format = output_format ,
           # 'knitrBootstrap::bootstrap_document' ,
           params = params.list ,
           envir = new.env() ,
           encoding = "UTF-8"
           )
         
    }

   if ( file.exists( output.filename )) browseURL( output.filename ) 


```

